---
title: "Data Exploration Project"
author: "Sean Bellevue"
date: "8/3/2022"
output: html_document
---

```{r message=FALSE, warning=FALSE}


library(tidyverse)
library(readr)
library(readxl)


```

#Pull the data 
```{r message=FALSE, warning=FALSE}

id_name_link <- read_csv("../Rawdata/id_name_link.csv")
Scorecard <- read_csv("../Rawdata/Most+Recent+Cohorts+(Scorecard+Elements).csv")
trends_1 <- read_csv("../Rawdata/trends_up_to_finish.csv")
trends_2 <- read_csv("../Rawdata/trends_up_to_inter_1.csv")
trends_3 <- read_csv("../Rawdata/trends_up_to_inter_2.csv")
trends_4 <- read_csv("../Rawdata/trends_up_to_inter_3.csv")
trends_5 <- read_csv("../Rawdata/trends_up_to_inter_4.csv")
trends_6 <- read_csv("../Rawdata/trends_up_to_inter_5.csv")
trends_7 <- read_csv("../Rawdata/trends_up_to_inter_6.csv")
trends_8 <- read_csv("../Rawdata/trends_up_to_UM.csv")
trends_9 <- read_csv("../Rawdata/trends_up_to_UPhoenix.csv")
trends_10 <- read_csv("../Rawdata/trends_up_to_UT.csv")
trends_11 <- read_csv("../Rawdata/trends_up_to_UTMB.csv")
trends_12 <- read_csv("../Rawdata/trends_up_to_Yorktowne.csv")

```

#Merge togther all of the trend files to clean them all at once
```{r}
Trends <- rbind(trends_1
                , trends_2
                , trends_3
                , trends_4
                , trends_5
                , trends_6
                , trends_7
                , trends_8
                , trends_9
                , trends_10
                , trends_11
                , trends_12
                )

summary(Trends)
```

#Remove any row with a null value, with incomplete data we can't compare or view trends as easily
```{r}

Trends <- Trends %>% 
  na.omit()

summary(Trends)
```

#Take the Monthorweek column and separate it into year and month columns. Then convert to M-YY for the starting date to better summarise the data to one month per college (schname) by sum of index [assuming that index is a snapshot value and not a trend so summing the values will still provide the same comparison of increase or decreased interest between time periods].

```{r}

Trends_mo <- Trends %>% 
  mutate(Y = substr(monthorweek,1,4)) %>% 
  mutate(M = substr(monthorweek,6,7)) %>% 
  mutate(M_YYYY = paste0(M, sep = '-', Y)) %>% 
  select(1,2,6,9)

Trends_mo <- Trends_mo %>% 
  group_by(schid, schname, M_YYYY) %>% 
  summarise(Ind_sum = sum(index))

```

#Remove duplicate schname from the id_name_link data to remove duplicate school names for later merging ease. Createing merge column with the scorecard

```{r}

unique_schname <- id_name_link %>%
  group_by(schname) %>% 
  filter(n() == 1)

unique_schname <- unique_schname %>% 
  mutate(merge_col = paste0(unitid, sep = ':', opeid))

Scorecard <- Scorecard %>% 
  mutate(merge_col = paste0(UNITID, sep = ':', OPEID))

```

#Merging Scorecard and unique schname to get schname for later merging with the tends and paring down to variables of interest. Then filtering for schools that grant predominantly bachelor's and with median income after 10 year income listed.
```{r}

score_schname <- merge(Scorecard, unique_schname, by = "merge_col", all = TRUE)

#Predominant bachlors = PREDDG == 3
score_schname_red <- score_schname %>% 
  select(126,5,122,11) %>% 
  filter(PREDDEG == 3) %>% 
  filter(`md_earn_wne_p10-REPORTED-EARNINGS` != 'NULL' & `md_earn_wne_p10-REPORTED-EARNINGS` != 'PrivacySuppressed')

score_schname_red$md_income <- as.numeric(score_schname_red$`md_earn_wne_p10-REPORTED-EARNINGS`)



```

#Per the data dictionary for the scorecard income is grouped by less than 35000 and above 75000 then between those two values.
```{r}
score_schname_red <- score_schname_red %>% 
  mutate(Income_Bracket = case_when(md_income < 35000 ~ 'Low',
                                    md_income > 75000 ~ 'High',
                                    TRUE ~ 'Mid')) %>% 
  select(1,2,5,6)
```


#merge the trend data and school scorecard filtered data. This produces many NA values where the schools have been filtered out of the scorecard. Removing those values to produce the final dataframe for analysis.
#Creating dummy variables for income bracket and pre or post scorecard
```{r}

final_merge <- merge(Trends_mo, score_schname_red, by.x = "schname", by.y = "schname", all.x = TRUE)

final_df <- final_merge %>% 
  na.omit()

final_df$M_YYYY <- as.Date(paste0(substr(final_df$M_YYYY,1,2), sep = '/', '01', sep = '/', substr(final_df$M_YYYY,4,7)), format = '%m/%d/%Y')

final_df <- final_df %>% 
  mutate(norm_ind = (Ind_sum - mean(Ind_sum))/sd(Ind_sum))

final_df <- final_df %>% 
  mutate(Timing = case_when(M_YYYY < '2015-09-01' ~ 'Pre',
                            M_YYYY >= '2015-09-01'~ 'Post')) %>% 
  mutate(High = case_when(Income_Bracket == 'High' ~ 1,
                          TRUE ~ 0)) %>% 
  mutate(Mid = case_when(Income_Bracket == 'Mid' ~ 1,
                          TRUE ~ 0)) %>% 
  mutate(Low = case_when(Income_Bracket == 'Low' ~ 1,
                          TRUE ~ 0)) %>% 
  mutate(Pre = case_when(M_YYYY < '2015-09-01' ~ 1,
                         TRUE ~ 0)) %>% 
  mutate(Post = case_when(M_YYYY > '2015-09-01' ~ 1,
                         TRUE ~ 0))

```

#Sperating out pre-scorecard dates and post-scorecard dates to compare the changes in index values between them.
```{r}

final_df_pre_sc <- final_df %>% 
  filter(M_YYYY < '2015-09-01')

final_df_post_sc <- final_df %>% 
  filter(M_YYYY > '2015-09-01')


```
#create a file for the final data set for analysis

```{r}
write_csv(final_df, file = 'prepared_data.csv')

```
